---
title: "Practice 3"
author: "Smit Patil"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

1. Download the data set for the tutorial.

2. Follow this tutorial on applying kNN to prostate cancer detection and implement all of the steps in an R Notebook. Make sure    to explain each step and what it does.

```{r}
#Importing data 
prc <- read.csv("prostate_cancer.csv")
str(prc)

#Creating a new column with values Benign, Malignant
prc$diagnosis <- factor(prc$diagnosis_result, levels = c("B", "M"), labels = c("Benign", "Malignant"))

#Creating a function for normalization
normalize <- function(x) 
{ 
  return ((x - min(x)) / (max(x) - min(x))) 
}

#Converting the data into normalized form
prc_n <- as.data.frame(lapply(prc[3:10], normalize))
summary(prc_n$radius)

#Splitting the data for training and testing
prc_train <- prc_n[1:65,]
prc_test <- prc_n[66:100,]

#Getting labels for trainig and testing
prc_train_labels <- prc[1:65, 2]
prc_test_labels <- prc[66:100, 2]

#Importing library "class" to perform kNN algorithm
library(class)
prc_test_pred <- knn(train = prc_train, test = prc_test,cl = prc_train_labels, k=10)

#Importing library "gmodels" to display te resulting output in a CrossTable
library(gmodels)
CrossTable(x = prc_test_labels, y = prc_test_pred, prop.chisq = FALSE)

#Calculating accuracy
True_Negative <- 7
True_Positive <- 16
Accuracy <- ((True_Negative + True_Positive)/35)*100
Accuracy
```

3. Try another kNN implementation using caret package. Compare the accuracy of the two implementations.

```{r}
library(lattice)
library(ggplot2)
library(data.table)
library(dplyr)
library(caret)

#Creating training dataset
training <- prc_train
training$diagnosis_result <- as.factor(prc_train_labels)

#Creating testing dataset
testing <- prc_test
testing$diagnosis_result <- as.factor(prc_test_labels)

#Assigning train control values to variables  
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
set.seed(882)

#Using train function from caret package to train the data
knn_fit <- train(diagnosis_result ~., data = training, method = "knn",
 trControl=trctrl,
 preProcess = c("center", "scale"),
 tuneLength = 10)

#Displaying accuracy of trained data and plotting the same
knn_fit
plot(knn_fit)

#testing the the model with testing data
test_pred <- predict(knn_fit, newdata = testing)
test_pred
```

4. Try the confusionMatrix function from the caret package to determine the accuracy of both algorithms.

```{r}
#Using ConfusionMatrix function from caret package for both models
confusionMatrix(test_pred, testing$diagnosis_result )
prc_test_labels_1 <- as.factor(prc_test_labels)
confusionMatrix(prc_test_pred, prc_test_labels_1)
```
