---
title: "Practicum 3"
author: "Smit Patil"
output: pdf_document
---

---Problem 2---

1. Download the data set Plant Disease Data Set. Note that the data file does not contain header names; you may wish to add those. The description of each column can be found in the data set explanation. This assignment must be completed within a separate R Markdown Notebook. Use read.transaction() from the arules package to read the data.

```{r warning = FALSE, message=FALSE}
#Importing libraries
library(arules)

#Importing plant data
plant.data <- read.transactions("plants.data")

#Summarising plant data
summary(plant.data)
```

2. Explore the data set as you see fit and that allows you to get a sense of the data and get comfortable with it. Is there distributional skew in any of the features? Is there a need to apply a transform?

```{r warning = FALSE, message=FALSE}
#Inspecting first 5 rows of the plant data
inspect(plant.data[1:5])

#Calculating frequency for the first 5 rows
itemFrequency(plant.data[,1:5])

#Plotting the ietms with atleast 10% support for first 5 species
itemFrequencyPlot(plant.data[1:5], support = 0.1)

#Plotting the top 10 species
itemFrequencyPlot(plant.data, topN = 10)

#Creating a sparse matrix for the first 10k species
image(plant.data[1:10000])

#Creating a sparse matrix for a random sample of 10k species
image(sample(plant.data, 10000))
```

3. Use association rules to segment the data similar to what was done in H채m채l채inen, W., & Nyk채nen, M. (2008, December). Efficient discovery of statistically significant association rules. In Data Mining, 2008. ICDM'08. Eighth IEEE International Conference on (pp. 203-212). IEEE.

```{r warning = FALSE, message=FALSE}
#Calculating rules using the default values of the apriori function
apriori(plant.data)

#Calculating rules using custom values
plant.rules <- apriori(plant.data, parameter = list(support = 0.0005, confidence = 0.9))

#summary of plant rules
summary(plant.rules)

#Inspecting the first 5 rules of the plant rules
inspect(plant.rules[1:5])

#Combining the sort function with the inspect function to find the best five rules according to the lift statistic
inspect(sort(plant.rules, by = "lift")[1:5])
```

4. Are there clusters in the data? Can plants be segmented into groups? Build a k-means clustering model to investigate. 

```{r warning = FALSE, message=FALSE}
library(klaR)
library(reshape2)

#Converting plant data to vector
plant.vector <- as.vector(unlist(plant.data@itemInfo))

#Saving the vector file as a .csv file
write(x = plant.vector, "plant.vector.csv", sep = ",")

#Reading the plant vector csv file as a data frame
plant.vector.data <- read.csv("plant.vector.csv", header = F, col.names = 1:100)

#Removing empty columns from the data
plant.vector.data <- plant.vector.data[,which(!is.na(plant.vector.data[1,]))]

#Looking at the head and the structure of the data
head(plant.vector.data)
str(plant.vector.data)

#Recasting the plant data
plant.new <- recast(plant.vector.data, X1 ~ value, id.var = "X1")

#Summary of the new plant data
summary(plant.new)

#Creating five clusters for plant data
plant.cluster <- kmeans(plant.new[,-1], 5)

#Printing the plant cluster sizes
plant.cluster$size

#Looking at the plant clusters along with their centers
plant.cluster$centers
```

5. Visualize the clusters.

```{r warning=FALSE, message=FALSE}
#Importing libraries
library(factoextra)
library(scatterplot3d)

#Plotting 2D graph to represent the clusters
fviz_cluster(plant.cluster, data = plant.new[,-1],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
             )
#Plotting scatter plot for first 3 values in 3D
scatterplot3d(plant.new[,-1], pch = plant.cluster$cluster, color = plant.cluster$cluster)
```
