---
title: "Practicum 3"
author: "Smit Patil"
output: pdf_document
---

---Problem 1---

1. Download the data set Bank Marketing Data Set. Note that the data file does not contain header names; you may wish to add those. The description of each column can be found in the data set explanation. Use the bank-additional-full.csv data set. Select an appropriate subset for testing. Use bank-additional.csv if your computer cannot process the full data set.

```{r warning = FALSE, message=FALSE}
#Importing bnak dataset
bank.data <- read.csv("bank-additional-full.csv", sep = ";", stringsAsFactors = T)

#Looking at the head and the structure of the bank data
head(bank.data)
str(bank.data)

# Summary of bank data
summary(bank.data)
```

2. Explore the data set as you see fit and that allows you to get a sense of the data and get comfortable with it. Is there distributional skew in any of the features? Is there a need to apply a transform?

```{r warning = FALSE, message=FALSE}
#Checking for any NA's in the dataset
anyNA(bank.data)

#Using for loop to print histograms of bank data
for(i in 1:ncol(bank.data))
  {
    sprintf("Histogram for: ", colnames(bank.data[i]))
    hist((as.numeric(bank.data[,i])), xlab = colnames(bank.data[i]), col = '#46495a', main = NULL)
 }

#Converting factor coloumns to numeric
for (i in 1:ncol(bank.data)-1)
  {
    if(is.factor(bank.data[,i]))
      {
        bank.data[,i] <- as.numeric(bank.data[,i])
      }
  }

#Printing the structure of bank data
str(bank.data)

#Creating a function for normalization
normalize <- function(x)
  {
    return ((x-min(x))/(max(x)-min(x)))
  }

#normalizing the bank data
bank.data <- data.frame(lapply(bank.data[,-21], normalize), bank.data$y)
```

3. Build a classification model using a support vector machine that predicts if a bank customer will open a term deposit account.

```{r warning = FALSE, message=FALSE}
#Importing libraries
library(caret)
library(e1071)

#Creating a random sample using createDataPartition function
set.seed(1)
sample <- createDataPartition(bank.data$bank.data.y, p=0.75, list = FALSE)

#Creating training and testing datasets SVM
bank.train.SVM <- bank.data[sample,]
bank.test.SVM <- bank.data[-sample,]

#Creating SVM model
bank.SVM <- svm(bank.data.y ~ ., data = bank.train.SVM, probability = TRUE)

#Predicting the values for the test data
SVM.predict <- predict(bank.SVM, bank.test.SVM, probability = TRUE)

#Printing the confusionMatrix for the SVM model
confusionMatrix(SVM.predict, bank.test.SVM$bank.data.y)
```

4. Build another classification model using a neural network that also predicts if a bank customer will open a term deposit account.

```{r warning = FALSE, message=FALSE}
#Importing libraries for Neural Netwrok
library(neuralnet)

#Creating training and testing datasets for Neural Netwrok(NN)
bank.train.NN <- bank.data[sample,]
bank.test.NN <- bank.data[-sample,]

#Setting y as an integer
bank.train.NN$bank.data.y <- as.integer(bank.train.NN$bank.data.y)
bank.test.NN$bank.data.y <- as.integer(bank.test.NN$bank.data.y)

#Creating function softplus for smoothening
softplus <- function(x) log(1+exp(x))

#Creating Neural Network model
bank.NN <- neuralnet(bank.data.y ~ ., bank.train.NN, hidden = c(3, 2), threshold = 0.5, rep = 1, linear.output = FALSE, act.fct = softplus)

#plotting the neural network graph
plot(bank.NN, rep="best")

#Computing the values for the test data
NN.predict <- compute(bank.NN, bank.test.NN[,-21])

#Calculating correlation between the actual and predicted values
cor.NN <- cor(NN.predict$net.result, bank.test.NN$bank.data.y)
sprintf("The correlation between actual values and predicted values by Neural Network is: %s", cor.NN)
```

5. Compare the accuracy of the two models based on AUC.

```{r warning = FALSE, message=FALSE}
#Importing libraries to calculate ROC and AUC
library(pROC)

#Retiving probability values for SVM
df <- as.data.frame(attr(SVM.predict, "probabilities"))
colnames(df) <- c("one", "two")

#Plotting AUC graph for SVM
plot.roc(roc(as.numeric(bank.test.SVM$bank.data.y), as.numeric(df$one)), axis = TRUE, legacy.axes = FALSE, print.auc = TRUE, asp = 0)

#Plotting AUC graph for Neural Network
plot.roc(roc(as.numeric(bank.test.NN$bank.data.y), as.numeric(NN.predict$net.result)), axis = TRUE, legacy.axes = FALSE, print.auc = TRUE, asp = 0)
```

6. Calculate precision and recall for both models. See this article to understand how to calculate these metrics.

```{r warning = FALSE, message=FALSE}
#Calculating precision for SVM
SVM.precision <- posPredValue(SVM.predict, bank.test.SVM$bank.data.y, positive = "no")
sprintf("The precision for SVM is: %s", SVM.precision)

#Calculating recall for SVM
SVM.recall <- sensitivity(SVM.predict, bank.test.SVM$bank.data.y, positive = "no")
sprintf("The recall for SVM is: %s", SVM.recall)

#Creating data frame prediction strength for Neural Netwrok based on the original factor values
NN.predict.strength <- ifelse(NN.predict$net.result>1.5, 2, 1)

#Calculating precision for Neural Network
NN.precision <- posPredValue(as.factor(NN.predict.strength), as.factor(bank.test.NN$bank.data.y), positive = "1")
sprintf("The precision for Neural Network is: %s", NN.precision)

#Calculating reacll for Neural Network
NN.recall <- sensitivity(as.factor(NN.predict.strength), as.factor(bank.test.NN$bank.data.y), positive = "1")
sprintf("The recall for Neural Network is: %s", NN.recall)
```
